Appendix of Bauer et al. (2023)

### 2023-02-15 manuscript_appendix.Rmd
Create manuscript_appendix.Rmd

---
title: "Development of restored dike grasslands to reference needs more than four years and can hardly modified by local site conditions"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  officedown::rdocx_document:
    reference_docx: template_word.docx
csl: template_references_j_appl_ecol.csl
bibliography: bibliography.bib
---

# Document head (remove later) <!--# Remove --> {.unnumbered}

@@

Restoration Ecology, TUM School of Life Sciences, Technical University of Munich, Germany

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

Max 8000 words <!--# Remove -->

# Abstract (max 300 words) {.unnumbered}

<!--# Write -->

@@

# Keywords {.unnumbered}

Bayesian models

dike grasslands

embankment

endangered species

environmental filter

species composition

restoration

sand

soil properties

surveys

vegetation

# Introduction

Sowing as an approach to restore semi-natural grasslands has long been established [@Kiehl.2010], is also able to restore functional diversity [@Engst.2016] and is continuously developed [e.g., @Freitag.2021]. Restoration is seen as a key factor to sustain ecosystem services and biodiversity [@COP.2014; @UnitedNations.2019b]. It is necessary to successfully restore grasslands since this habitat type provide several ecosystem services and a high biodiversity while it is worldwide endangered [@Dengler.2014; @Bardgett.2021] even by restoration through unsuited afforestation [@dudley2020]. In Europe, calcareous grasslands and hay meadows are on the red list [Category 3, vulnerable, @Janssen.2016].

Sowing high-diversity seed mixtures of local provenance produced by commercial but specialized seed companies is a promising way to scale up restoration efforts. Though, there are still open questions about specifying the seed mixtures to specific site conditions or even more to future climate conditions to improve efficiency and effectiveness of restoration efforts [@Torok.2021]. For this reason, restoration ecology has to increase predictability of restoration approaches [@Suding.2011; @Mouquet.2015]. Rigorous, repeatable, and transparent experiments linked to a hypothesis derived from theory are needed to guide evidence-based restoration [@Cooke.2018; @Wainwright.2018]. Local site conditions and the restoration approach are key predictors of the development of vegetation after sowing [@Brudvig.2017b]. Seeding in general is a way to overcome the dispersal filter [@myers2009; @orrock2023], but we focus on environmental and the biotic filtering which are the main assembly processes to be manipulated by the choice of the right seed mixture-substrate combination [@Torok.2017]. This means a mutually adaption of the substrate to the preferences of the target vegetation, and the seed mixtures to the characteristics of the chosen substrate. The right substrates should reduce habitat filtering and the right seed mixture should minimize competitive exclusion of unsuited species. Modifying seed mixtures for certain site conditions came up with functional plant traits [@Funk.2008; @Laughlin.2014; @Balazs.2020] but it is not easy to put it into practice for novel ecosystems like urban grasslands [@Bauer.2022]. Substrates cannot only be modified for conservation purposes in urban areas, but also in quarries [@chenot-lescure2022] or dikes [@Liebrand.1996].

Dikes are promising sites for the restoration of species-rich grasslands [@Liebrand.1996], because they can increase habitat area and connectivity of semi-natural grasslands and therefore significantly contribute to biodiversity conservation in an agricultural landscape [@Batori.2020]. Dikes have specific characteristics with strongly exposed slopes and river and land orientation [@Batori.2016] and need a reliable vegetation cover to conserve dike security [@KleberLerchbaumer.2017]. Dikes can reconcile several requirements [@teixeira2022], but the seed mixture-substrate combination must be adapted.

The aim of this study is to identify the ideal combination of substrates and seed mixtures for north and south exposed slopes in Central Europe to develop a vital and species-rich vegetation cover. We set up an experiment and tested agricultural soil with different amounts of sand admixture and different installation depths, as well as different seed mixture types and different seed densities. We expected a better development of dry grassland on the south exposition with sandy substrate and hay meadows on north exposition with less sandy substrate. Since seeds are expensive, practitioners like to know if the slopes need a higher seed density [cf. @KleberLerchbaumer.2017].

To measure the success of restoration, we chose four indices. First, we observed the persistence of the sown species, which is a key interest of seeding as a restoration approach [@Wilsey.2021]. For this, it was surveyed if the sown species are present in the respective plot in following years. Second, we calculated the Favorable Conservation Status (FCS) which distinguishes characteristic diversity and derived diversity [@Helm.2015]. Characteristic diversity is created by established species which belong to the habitat-specific species pool. In other words, these species are target species for a habitat type of interest. Third, the successional distance to the reference was determined. This measure quantifies the distance of a plot to the average reference site from an ordination [@Rydgren.2019]. To complement the indices, the success of restoration or the difference from desired conditions needs to be evaluated [@Brudvig.2017b]. Therefore, the species composition of the experimental plots were compared with reference sites. We asked following question to evaluate the restoration approaches and success after four years:

1.  How strong differs the persistence of the sown species among the seed mixture-substrate combinations?

2.  How strong differs the Favorable Conservation Status (FCS) among the seed mixture-substrate combinations?

3.  How close is the vegetation to the reference state?

# Material & Methods

## Experimental design {.unnumbered}

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:map)); 314 m above sea level; WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and an annual precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@hari2020; @cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix Table @ref(tab:flood-weather), @lfu2021waterlevel; @lfu2021discharge]. Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from removed agricultural soil of a dike construction site near Steinkirchen. The substrates were mixed with a big roller and an excavator prepared the plots. The target vegetation were lowland hay meadows [EUNIS code: R22, @Chytry.2020; Arrhenatherion elatioris, EuroVegChecklist: CM01A, @mucina2016] and semi-dry calcareous grassland (R1A; Cirsio-Brachypodion pinnati, DA01B). The species pool of hay meadows and dry grasslands consisted of 55 and 58 species, respectively. The seeds were received from a commercial producer of autochthonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Figure \@ref(fig:traits)). As expected, the hay meadows mixtures have a higher specific leaf area (SLA) lower seed mass and higher canopy height than the dry meadows mixtures. The south exposed plots were sown in mid-April 2018 and the north exposition two weeks later. Due to the drought, the south exposition was protected by a geotextile consisting of straw chaff (350 g m^-2^) which was removed after two weeks.

The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, *Bromus hordeaceus* was sown as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition, halfway up the slopes of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three sand admixtures (0%, 25% and 50%) and tow soil depths (15 vs. 30 cm) of practical interest [@KleberLerchbaumer.2017] and two . Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. We tested soil samples of the three substrates on both expositions by mixing several sub-samples. The sand admixture changed the soil texture, increased the C/N ratio, but reduced Calciumcarbonat, and hardly changed the rather high pH values (Table \@ref(tab:substrate)). Phosphate and potassium were rather scarce for agricultural soils and Magnesium rather high [@lfl2022]. The two seed mixture types were tested with the normal seeding density and the increased density recommended for slopes [4 vs. 8 g m^-2^ @Kiehl.2010; @KleberLerchbaumer.2017]. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species (Appendix Table \@ref(tab:establishment)).

## External data

Reference vegetation surveys were extracted from sPlotOpen [@Sabatini.2021] and our own surveys in the surroundings on Danube dikes [@bauer2023]. We selected six calcareous grassland plots [EUNIS code R1A, @Chytry.2020] within SE Germany from sPlotOpen and 98 plots of our own survey which included furthermore lowland hay meadows (R22) and as a negative reference ruderal, dry and anthropogenic vegetation (V38). The functional traits of specific leaf area (SLA; available traits: 99.2% of all species), seed mass (98.8%) and canopy height (99.6%) were retrieved from the databases LEDA [@Kleyer.2008] and TRY [@Kattge.2020].

## Statistical analyses {.unnumbered}

To measure the effects of the treatments on our three response variables, we calculated Bayesian linear mixed-effects models (BLMM) with the random effect plot nested in block with the Cauchy prior [see @lemoine2019]. Furthermore, we included as a fixed effect the botanists of each plot. For the simple effects of the treatments, we chose plausible weakly informative priors. To evaluate the influence of the priors, prior predictive checks and models with non-informative priors were calculated. For the computation, we used 4 chains, a thinning rate of 2, 5000 iterations for warm-up and 10,000 in total. For evaluating the computation, MCMC chain convergence was checked with trace plots and by the R-hat values and MCMC chain resolution by the effective sampling size (ESS). Posterior predictive check was done with Kernel density estimates histograms of statistics skew and leave-one-out (LOO) cross-validation [see @gabry2019]. Finally, models were compared with the Bayes factor (BF) and Bayesian *R*² values [@gelman2019].

A non-metric multidimensional scaling ordination (NMDS) with Jaccard dissimilarity was used to visualize variation in species composition in space and time. Seven species which had an accumulated cover over all plots of less than 0.5% were excluded (2D-stress 0.21).

Data, code and the whole model specification and evaluation is stored on Zenodo and in an easily accessible a scroll-down document on GitHub [!cite Zenodo!]<!--# Cite Zenodo--> There, the sections are referenced to the Bayesian analysis reporting guidelines [BARG, @kruschke2021]. All analyses were performed in R [Version 4.2.2, @base] with the functions 'brm' from the package 'brms' [@brms] for model calculation, several functions from 'brms' and 'bayesplot' [@bayesplot] for model evaluation, 'metaMDS' from 'vegan' for the ordination [@vegan] and 'dbFD' from 'FD' for the calculation of community-weighted means of functional traits [@FD].

# Results

We could identify a statistically clear positive effect of sand addition on the persistence of sown species and on the recovery rate (Figure \@ref(fig:effects)). Furthermore, sown hay meadows were clearly closer to their reference plots than sown dry grasslands were to their references. The posterior distributions which are shown in the interactions plots which separates exposition and survey year indicate for all three responses a strong positive development after one year with slow increase in following years (Figures \@ref(fig:persistence); \@ref(fig:fcs); \@ref(fig:recovery)) . Both expositions have similar trends but for all responses the values are clearly lower on the south exposition, e.g., persistence values are on average more than 46% higher on the north exposition (Figure \@ref(fig:persistence)). The effect of restoration treatments is in the interaction plots not clear. Hay meadow seed mixture led to a closer development to hay meadow references and almost reached on the north exposition the area of the standard deviation of reference sites (Figure \@ref(fig:recovery)). Though, the sand admixture had no clear effect on the development. Persistence of both seed mixture types seems to be slightly positive affected by sand admixture on the north exposition (+ 6--7 ± 4%, Figure \@ref(fig:persistence)).

The ordination visualizes the species composition of seed mixtures and the development of the plots over time (Figure \@ref(fig:nmds)). The NMDS confirms that the seed mixtures are variable, but distinctive for hay meadows and dry grasslands. In addition, it visualizes the decreasing recovery rate on the NMDS1 axis which was already presented (Figure \@ref(fig:recovery)) but the right direction of the development. One exception is that the development of intended hay meadows on the south exposition does not develop towards its seed mixture compositions. The ordination presents the experimental plots not only in context with seed mixtures, but also with reference sites. The reference sites have a bigger variation than the seed mixtures and are close to the seed mixtures but hardly overlap. The positions of reference sites are shifted to the left on the NMDS1 axis which means in the direction of early successional states. On the south exposition, we had negative reference sites which are located in the NMDS in between the positive reference sites and the recent state of restored plots. However, most of negative reference sites (ruderal, dry and anthropogenic vegetation) are situated in the upper half and explicitly recent restoration sites on the south exposition are close to them.

# Discussion

The vegetation of the dike experiment developed in the right direction, but with reduced velocity and significantly and lasting worse development of the south exposition. Sand admixture of 25% was sufficient to reach a positive effect. All in all, it was hardly possible to identify an ideal seed mixture-substrate combination. Though, hay meadows developed closer to its reference on the north exposition than dry grasslands.

## General effects of treatments and exposition

As expected, sand admixture led to higher persistence of sown species but only on the north exposition and the effect size of about 6% in the fourth year after restoration is rather small. In addition, 50% admixture did not further improve the result. Restoration on agricultural soils can have limited success due to high nutrient loads [@walker2004] but can lead to positive results, if mixed with mineral material as sand and gravel [@chenot-lescure2022]. Substrate depth did not significantly affect persistence or Favorable Conservation Status (FCS) similarly to earlier studies [@baer2004]. Larger differences of soil depth might be necessary to create negative effects by greater soil depths as was shown for prairies [@dornbush2010]. Neither, seeding density had an effect on persistence and FCS which fits to result of @Kaulfuss.2022, who found out that a certain amount of seeds are necessary but higher densities do not further improve establishment of target species but have rather a slight negative effect.

The vegetation on the south exposition performed consistently and significantly worse than the north exposition and had a different species composition which was also shown by @Batori.2016 in Hungary. Here, the differences might have some unwanted experimental reasons: the geotextile, which had been implemented on the south exposition, was removed after two weeks because of the death of several protected lizards. This led to the drying of many seedlings why the distinction of the seed mixture types might be leveled on the south exposition. Moreover, the intense drought of 2018 and 2019 [@hari2020] might have decreased germination and establishment of sown species because a moist spring with a mesic temperature improves establishment [@Larson.2021; @orrock2023]. The interesting point here is the lasting negative effect which suggests a legacy effect by unfavorable weather conditions after sowing as it was observed by other studies [@Stuble.2017; @Groves.2020]. The problems during the establishment phase might have led to an alternative trajectory and could probably only be met by resowing.

## No interaction effect of seed mixture-substrate combinations

Our aim was to identify a ideal seed mixture-substrate combination in respect of restoration effectiveness and biodiversity. For evaluating effectiveness, we measured the establishment and persistence of the sown species and for fostering plant biodiversity in general, we measured the FCS. Though, we could not identify an interaction effect for neither of the indices. We would have expected a better performance of hay meadow seed mixtures with lower sand admixture and for dry grasslands with higher sand admixture. Our results suggest that, at least after four years, the substrate conditions are within the range of both seed mixture types (hay meadows, dry grasslands). Although, both types are clearly plant-sociologically and functionally distinct, they are still relative closely related why they contain shared species and could develop on similar site conditions [Figure @ref(fig:traits); @Oberdorfer.1993; @Oberdorfer.1993b]. Other grassland studies could identify more or less clear interactions of opposing habitat preferences or functional traits along the gradients of productivity, moisture, and nutrients [@Zirbel.2020; @Freitag.2021; @Kaulfuss.2022]. Though, these two studies did not work with an experimental set up of different seed mixture-substrate combinations, but analyzed the result of habitat and biotic filtering after five and about 15 years, respectively. Furthermore, the non-existence of an ideal combination could be explained by priority effects which means that the species of the sown species could establish earlier than others and preempted the available niches [@Fukami.2015].

## The success of the restoration approaches

The seed mixtures and positive references were similar but hardly overlapped (Figure \@ref(fig:nmds)). The position on the NMDS suggests that the seed mixture represent a latter successional stage than the positive references. The NMDS shows a slightly better adaption of hay meadows to the north exposition than of dry grasslands (Figure @ref(fig:recovery)). This can be expected from the requirements of hay meadows for mesic conditions which can be provided on north exposed dike slopes [@Batori.2020; @Oberdorfer.1993b]. On the south exposition, the plots of hay meadows developed rather to dry grassland references which indicates an inefficient restoration by using an non-adapted seed mixture.

The vegetation developed in general in the right direction but was still distinct from positive references and seed mixtures after four years. It was rather similar to the negative reference of perennial ruderal vegetation. The gap between goal and restoration outcome was also shown for other seeding experiments or restorations [@mitchley2012; @Engst.2016; @Kaulfuss.2022] or for dike vegetation compared with semi-natural references [@Batori.2016]. This result is not surprising since the 'recovery debt' is a general phenomenon of grassland restoration [@MorenoMateos.2017; @jones2018].

# Conclusions

Our results suggest that adapted seed mixtures can increase restoration efficiency by seeding hay meadows on the north exposition and not on the south exposition of dikes. Furthermore, to reduce nutrient load by sand admixture has a clear positive effect, although the effect is small. The question remains if money is best invested in sand admixture or in other restoration measures. It seems not to be necessary to increase seeding density on dike slopes, which would reduce restoration costs. As well as the use of more deeper soil depth which would otherwise create costs for removal of topsoil. There was no perfect seed mixture-substrate combination, why we conclude that a variation of seed mixtures types and different soils along restoration sections would promote biodiversity better than a single solution [@bauer2023; @Holl.2022]. The strong and lasting negative effect of drought on the south exposition might require a resowing. To close the recovery debt, the adaption of the management might be promising since it is a relevant factor aside of restoration approach and site characteristics [@Grman.2013; @tölgyesi2021]. For example, starting sheep grazing on the restored sites will modify the disturbance regime and improve dispersal. All in all, our results support the finding that dike grasslands can promote biodiversity in agricultural landscapes [@Batori.2020]. Though, the recovery debt highlights that restored grasslands cannot substitute old-growth grasslands [@nerlekar2020; @Buisson.2022].

# Acknowledgements {.unnumbered}

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) and Klaus Rachl and Stefan Radlmair (Regierung von Niederbayern) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship. The study has been supported by the TRY initiative on plant traits ([http://www.try-db.org).](http://www.try-db.org).) The TRY initiative and database is hosted, developed and maintained by J. Kattge and G. Boenisch (Max Planck Institute for Biogeochemistry, Jena, Germany). TRY is currently supported by Future Earth/bioDISCOVERY and the German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig.

# Author contribution {.unnumbered}

JH and JK designed the experiment. JH did the surveys in the years 2018--2020 and MB in 2019 and 2021. MB did the analyses and wrote the manuscript. JK and JH critically revised the manuscript.

# Open research {#sec-open-research .unnumbered}

Data and code are stored on Zenodo (Bauer et al. 2023) <!--# Update citation after upload -->

Model evaluation is stored on GitHub: [github.com/markus1bauer/2023_danube_dike_experiment](https://github.com/markus1bauer/2023_danube_dike_experiment/tree/main/markdown)

# Funding {.unnumbered}

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020 (No. 80 002 312).

# References {.unnumbered}

::: {#refs}
:::

# Tables {.unnumbered}

## Table 1 {.unnumbered}

```{r tab.id = "seedmix", tab.cap = "Each plot received an individual set of 20 species with some restrictions to the number of species per functional group. The total species pool for hay meadows was 55 and for dry grassland 58. In total, seed mixtures had 16 or 32 g per plot. All individual seed mixtures are stored in Table A3."}
read_csv(here::here("data", "raw", "data_raw_table_1_seedmix.csv"),
         show_col_types = FALSE) %>%
  slice(-1) %>%
  flextable() %>%
  merge_at(part = "header", j = 2:3) %>%
  align(align = "center", part = "all", j = -1) %>%
  add_header_row(top = FALSE, values = c("", "Hay meadow", "Dry grassland", "", "", "")) %>%
  add_header_row(top = FALSE, values = c("", "#", "#", "#", "%", "%")) %>%
  bold(part = "header", i = 1) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  autofit()
```

## Table 2 {.unnumbered}

```{r tab.id = "substrate", ft.align = "left", tab.cap = "Soil of the three substrates was analyzed for the fraction < 2mm fraction. Soil texture was classified according to the 'Bodenkundliche Kartieranleitung' [@BodenkundlKartieranl.2005]. The pH was measured in CaCl2 solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in CaCl2 extract. For calculating CaCO3, a subsample was annealed at 550 °C and the measured C amount calculated with 8.33. To calculate total N and the C/N ratio, a subsample was incinerated at 1000 °C. Lt3 = medium clayey loam; Ls4 = strong sandy loam; Sl3 = medium loamy sand; Sl4 = strong loamy sand."}
read_csv(here::here("data", "raw", "data_raw_table_3_substrate.csv"),
         show_col_types = FALSE) %>%
  select(-starts_with("lfl")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(
      here::here("data", "raw", "data_raw_table_3_substrate.csv"),
                 show_col_types = FALSE) %>%
      select(-starts_with("lfl")) %>%
      slice(1)
    ) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  align(align = "center", part = "all", j = -1) %>%
  compose(i = 1, j = 10, part = "header", value = as_paragraph("P", as_sub("2"), "O", as_sub("5"))) %>%
  compose(i = 1, j = 11, part = "header", value = as_paragraph("K", as_sub("2"), "O")) %>%
  compose(i = 1, j = 12, part = "header", value = as_paragraph("Mg", as_sup("2+"))) %>%
  compose(i = 1, j = 14, part = "header", value = as_paragraph("CaCO", as_sub("3"))) %>%
  compose(i = 2, j = 10, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  compose(i = 2, j = 11, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  compose(i = 2, j = 12, part = "header", value = as_paragraph("mg 100 g", as_sup("−1"))) %>%
  align(align = "left", part = "footer") %>%
  set_table_properties(layout = "autofit", width = 1)
```

# Figures {.unnumbered}

## Figure 1 {.unnumbered}

```{r fig.id = "map", fig.asp = 1.2854, fig.width = 5, fig.cap = "The experiment was located on a dike at river Danube in SE Germany.  The 288 plots were allocated in six blocks (red squares on upper photo) and on the north and south exposition [central photo; both aerial photos: @bayerischevermessungsverwaltung2023]. Four treatments were conducted (Sand admixture, substrate depth, seeding density, seed mixture types H and D). The western half of a block was with a shallow substrate depth for the blocks and within this half the substrates. The photo on the bottom shows one block from the side after four years (Photo: Markus Bauer)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_1_map.png")
  )
```

## Figure 2 {.unnumbered}

```{r fig.id = "traits", fig.asp = 0.3030, fig.width = 6.49, fig.cap = "The community-weighted means (CWM) of the seed mixtures used for the field experiment of the functional traits specific leaf area (SLA), seed mass and canopy height. The CWMs were weighted by seed weight in the seed mixture and  indicate differences between the target types hay meadows and dry grasslands . "}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm_traits.tiff")
  )
```

## Figure 3

```{r fig.id = "effects", fig.asp = 0.3030, fig.width = 6.49, fig.cap = "The posterior density distributions (grey) of the treatments calculated over all four surveyed years. Furthermore, the median and the 66% and 95% credible intervals are shown. We evaluated the persistence of sown species (A), the Favorable Conservation Status (FCS; B) and the successional distance to reference (recovery; C). The FCS is the ratio of target species and non-target species. Note that the zero lines indicate that both levels have equal values. This means that hay meadows are closer to their reference than dry grasslands  (C)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff")
  )
```

## Figure 4 {.unnumbered}

```{r fig.id = "persistence", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "Persistence of sown species on a dike experiment over the years. The substrates had different sand admixtures and different seed mixture types were sown. Per plot were 20 species sown. Shown are the posterior distributions, their medians and 95 % credible intervals derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x8cm.tiff")
  )
```

## Figure 5 {.unnumbered}

```{r fig.id = "fcs", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The Favorable Conservation Status (FCS) of a dike experiment. The substrates had different sand admixtures and different seed mixture types were sown. The zero line indicate that target and non-target species are balanced. Positive values indicate that there are more target species.  Shown are the posterior distributions, their medians and 95 % credible intervals derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x8cm.tiff")
  )
```

## Figure 6 {.unnumbered}

```{r fig.id = "recovery", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The successional distance (*d~jt,0~*) of plots of an experiment on a dike to its respective reference sites (hay meadows or dry grasslands). The substrates had different sand admixtures and different seed mixture types were sown. The zero lines indicate the mean position of the reference sites on the NMDS axis 1 (Figure 7). The grey area marks the standard deviation of the position of the reference sites. Shown are the posterior distributions, their medians and 95 % credible intervals derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x8cm.tiff")
  )
```

## Figure 7 {.unnumbered}

```{r fig.id = "nmds", fig.asp = 0.9696, fig.width = 6.49, fig.cap = "NMDS ordination based on Sørensen dissimilarity of the species composition of 288 plots on dike grasslands over four years after sowing the seed mixtures (black squares).  In addition, 98 positive and negative reference plots were included (filled symbols) from older dike grasslands in the surroundings [@bauer2023] and six plots from sPlotOpen [@Sabatini.2021].  The circles show the standard error of the groups. 2D stress: 0.21."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff")
  )
```

## Table A1 <!--# Is it cited in the text? --> {.unnumbered}

## Table A2 <!--# Is it cited in the text? --> {.unnumbered}

@@

## Session Info {.unnumbered}

```{r sessioninfo}
sessionInfo()
```

>>>>

---
title: "Appendix of Bauer et al. (2023)"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  officedown::rdocx_document:
    reference_docx: template_word.docx
csl: template_references_j_appl_ecol.csl
bibliography: bibliography.bib
---

@@

## Table A1 {.unnumbered}

@@

## Table A2 {.unnumbered}

@@
